\Rpackage{MSnbase}: labelled MS2 data pre-processing, visualisation and quantification.

### 2012-09-07 inst/doc/MSnbase-demo/MSnbase-demo.Rnw

<<environment, cache=FALSE, echo=FALSE>>=
suppressPackageStartupMessages(library("MSnbase"))
suppressPackageStartupMessages(library("zoo"))
suppressPackageStartupMessages(library("doMC"))
suppressPackageStartupMessages(library("multicore"))
library("grid")
library("reshape2")
@

@@

<<readdata, echo=TRUE, cache=FALSE>>=
file <- dir(system.file(package = "MSnbase", dir = "extdata"),
            full.names = TRUE, pattern = "mzXML$")
rawdata <- readMSData(file, msLevel = 2, verbose = FALSE)
@

@@

<<MSnExp,cache=FALSE,echo=TRUE>>=
library("MSnbase")
itraqdata
head(fData(itraqdata))
@

<<experiment-size,echo=FALSE,cache=FALSE>>=
sz <- sum(sapply(assayData(itraqdata),object.size)) + 
  object.size(itraqdata)
sz <- round(sz/(1024^2),2)
@

@@

<<Spectrum,cache=FALSE,echo=TRUE>>=
sp <- itraqdata[["X1"]]
sp
@

@@

<<accessors,cache=FALSE,echo=TRUE>>=
peaksCount(sp)
head(peaksCount(itraqdata))
rtime(sp)
head(rtime(itraqdata))
@

@@

\begin{figure}[!ht]
<<spectrum-plot, dev='pdf', fig.width=4, fig.height=4, fig.keep='high'>>=
plot(sp, reporters = iTRAQ4, full = TRUE) 
@ 
\caption{Raw MS2 spectrum with details about reporter ions.}
\label{fig:spectrum-plot}
\end{figure}

<<bsasel, echo=FALSE>>=
bsasel <- fData(itraqdata)$ProteinAccession == "BSA"
@

@@

<<subset,echo=TRUE>>=
sel <- fData(itraqdata)$ProteinAccession == "BSA"
bsa <- itraqdata[sel]
bsa
as.character(fData(bsa)$ProteinAccession)
@

@@

\begin{figure}[p]
<<msnexp-plot, dev='pdf', fig.width=3.7, fig.height=4.5, fig.keep='last'>>=
plot(bsa, reporters = iTRAQ4, full = FALSE) + theme_gray(8)
@ 
\caption{Experiment-wide raw MS2 spectra. The y-axes of the individual 
  spectra are automatically rescaled to the same range. 
  See section \ref{sec:norm} to rescale peaks identically. }
\label{fig:msnexp-plot}
\end{figure}

@@

<<removePeaks,echo=TRUE,cache=FALSE>>=
experiment <- removePeaks(itraqdata, t = 400, verbose = FALSE)
## total ion current
ionCount(itraqdata[["X55"]])
ionCount(experiment[["X55"]])
@

@@

<<preprosp, cache=FALSE, echo=FALSE>>=
int <- c(0,1,1,3,1,1,0,0,0,1,3,7,3,1,0)
mz <- c(113.9,114.0,114.05,114.1,114.15,114.2,114.25,
        114.3,114.35,114.4,114.42,114.48,114.5,114.55,114.6)
ppsp <- new("Spectrum2",intensity=int,mz=mz,centroided=FALSE)
p1 <- plot(ppsp, full = TRUE, plot = FALSE) + theme_gray(5) + 
  geom_point(size=3,alpha=I(1/3)) +
  geom_hline(yintercept=3,linetype=2) + 
  opts(title="Original spectrum") 
p2 <- plot(removePeaks(ppsp,t=3), full=TRUE, plot = FALSE) + 
  theme_gray(5) +
  geom_point(size=3,alpha=I(1/3)) + 
  geom_hline(yintercept=3,linetype=2) + 
  opts(title="Peaks < 3 removed") 
p3 <- plot(clean(removePeaks(ppsp,t=3)), full = TRUE, plot = FALSE) + 
  theme_gray(5) +
  geom_point(size=3,alpha=I(1/3)) + 
  geom_hline(yintercept=3,linetype=2) + 
  opts(title="Peaks < 3 removed and cleaned") 
@

\begin{figure}[p]
<<preproc-plot, dev='pdf', fig.width=3, fig.height=6, echo=FALSE>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(3, 1)))
print(p1, vp=vplayout(1,1))
print(p2, vp=vplayout(2,1))
print(p3, vp=vplayout(3,1))
@ 
\caption{This figure illustrated the effect or the \Rfunction{removePeaks} 
  and \Rfunction{clean} methods. The left-most spectrum displays two peaks, 
  of max height 3 and 7 respectively. The middle spectrum shows the result 
  of calling \Rfunction{removePeaks} with argument \Robject{t=3}, which sets 
  all data points of the first peak, whose maximum height is smaller or equal 
  to \Robject{t} to 0. The second peak is unaffected. Calling \Rfunction{clean} 
  after \Rfunction{removePeaks} effectively deletes successive 0 intensities from 
  the spectrum, as shown on the right plot. }
\label{fig:preproc}
\end{figure}

@@

<<trimMz,echo=TRUE,cache=FALSE>>=
range(mz(itraqdata[["X55"]]))
experiment <- trimMz(experiment, mzlim = c(112,120))
range(mz(experiment[["X55"]]))
experiment
@ 
As can be seen above, all processing performed on the experiment is recorded 
and displayed as integral part of the experiment object.

@@

<<reporters,echo=TRUE,cache=FALSE>>=
mz(iTRAQ4)
width(iTRAQ4)
@

@@

\begin{figure}[p]
<<quantitation-plot, dev='pdf',fig.width=5, fig.height=5, echo=FALSE>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(2, 2)))
vplayout <- function(x, y)
  viewport(layout.pos.row = x, layout.pos.col = y)
print(p + opts(title = "Quantitation using 'sum'") +
      geom_point(size = 3, alpha = I(1/3), colour = "red"),
      vp = vplayout(1, 1))
print(p + opts(title = "Quantitation using 'max'") +
      geom_point(aes(x = 114.1, y = 3),
                 alpha = I(1/18),
                 colour = "red", size = 3),
      vp = vplayout(1, 2))
print(p +
      opts(title = "Trapezoidation and strict=FALSE") +
      geom_polygon(alpha = I(1/5), fill = "red"),
      vp = vplayout(2, 1))
print(p +
      opts(title = "Trapezoidation and strict=TRUE") + 
      geom_polygon(aes(x = c(NA, 114.05, 114.05, 114.1, 114.15, 114.15, NA),
                       y=c(NA, 0, 1, 3, 1, 0, NA)), fill = "red", alpha = I(1/5)),
      vp = vplayout(2,2))

@ 
\caption{The different quantitation methods are illustrated above. 
  Quantitation using \Robject{sum} sums all the data points in the peaks to 
  produce, for this example, \Sexpr{quantify(ssp,iTRAQ4[1],method="sum")[[1]]}, 
  whereas method \Robject{max} only uses the peak's maximum intensity, 
  \Sexpr{quantify(ssp,iTRAQ4[1],method="max")[[1]]}. 
  \Robject{Trapezoidation} calculates the area under the peak taking 
  the full with into account (using \Robject{strict=FALSE} gives
  \Sexpr{round(quantify(ssp,iTRAQ4[1],method="trap",strict=FALSE)[[1]],3)}) 
  or only the width as defined by the reporter 
  (using \Robject{strict=TRUE} gives
  \Sexpr{round(quantify(ssp,iTRAQ4[1],method="trap",strict=TRUE)[[1]],3)}). }
\label{fig:quant-methods}
\end{figure}

@@

<<quantify,echo=TRUE,cache=FALSE,keep.source=T>>=
qnt <- quantify(experiment,
                method = "trap",
                reporters = iTRAQ4,
                strict = FALSE,
                parallel = FALSE,
                verbose = FALSE)
qnt
head(exprs(qnt))
@

@@

<<removeNa,echo=TRUE>>=
table(is.na(qnt))
whichRow <- which(is.na((qnt))) %% nrow(qnt)
qnt <- qnt[-whichRow, ]
sum(is.na(exprs(qnt)))
@

@@

<<purityCorrect,echo=TRUE,cache=FALSE>>=
impurities <- matrix(c(0.929, 0.059, 0.002, 0.000,
                       0.020, 0.923, 0.056, 0.001,
                       0.000, 0.030, 0.924, 0.045,
                       0.000, 0.001, 0.040, 0.923),
                     nrow = 4)
qnt.crct <- purityCorrect(qnt, impurities)
head(exprs(qnt))
head(exprs(qnt.crct))
@

@@

<<normalise,echo=TRUE,cache=FALSE>>=
qnt.max <- normalise(qnt, "max")
qnt.sum <- normalise(qnt, "sum")
qnt.quant <- normalise(qnt, "quantiles")
qnt.qrob <- normalise(qnt, "quantiles.robust")
qnt.vsn <- normalise(qnt, "vsn")
@

@@

\begin{figure}[p]
<<norm-plot, dev='pdf', fig.width=5, fig.height=7, echo=FALSE>>=
.plot <- function(x,ttl=NULL) {
  boxplot(exprs(x),
          main=ifelse(is.null(ttl),processingData(x)@processing[2],ttl),
          cex.main=.8,
          cex.lab=.5,
          cex.axis=.5,
          cex=.8)
  grid()
}
oldmar <- par()$mar
par(mfrow=c(3,2),mar=c(2.9,2.9,2.9,1))
.plot(qnt, ttl = "Non-normalised data")
.plot(qnt.max, ttl = "Maximum")
.plot(qnt.sum, ttl = "Sum")
.plot(qnt.quant, ttl = "Quantile")
.plot(qnt.qrob, ttl = "Robust quantile")
.plot(qnt.vsn, ttl = "vsn")
@ 
\caption{ Comparison of the normalisation \Robject{MSnSet} methods. 
  Note that vsn also glog-transforms the intensities. }
\label{fig:norm-plot}
\end{figure}

<<cvdata,echo=FALSE,cache=FALSE>>=
sd1 <- apply(log2(exprs(qnt))+10,1,sd)
mn1 <- apply(log2(exprs(qnt))+10,1,mean)
cv1 <- sd1/mn1
sd2 <- apply(exprs(qnt.vsn)+10,1,sd)
mn2 <- apply(exprs(qnt.vsn)+10,1,mean)
cv2 <- sd2/mn2
dfr <- rbind(data.frame(rank=order(mn1),cv=cv1,norm="raw"),
             data.frame(rank=order(mn2),cv=cv2,norm="vsn"))
library("zoo")
## rmed1 <- rollapply(cv1,7,function(x) median(x,na.rm=TRUE))
## rmed2 <- rollapply(cv2,7,function(x) median(x,na.rm=TRUE))
##
## Calling directly rollapply.zoo to make it zoo_1.6-4 compatible.
## The above requires zoo >= 1.7-0, which is as of 15 March 2011
## not yet available on CRAN (only on r-forge).
rmed1 <- zoo:::rollapply.zoo(cv1,7,function(x) median(x,na.rm=TRUE))
rmed2 <- zoo:::rollapply.zoo(cv2,7,function(x) median(x,na.rm=TRUE))
dfr2 <- 
  rbind(data.frame(x=seq(1,70,by=(70/length(rmed1))),
                   y=rmed1,norm="raw"),
        data.frame(x=seq(1,70,by=(70/length(rmed1))),
                   y=rmed2,norm="vsn"))

@@

\begin{figure}[!ht]
<<cv-plot, dev='pdf', fig.width=5, fig.height=4, echo=FALSE>>=
print(p)
@ 
\caption{CV versus signal intensity comparison for log2 and vsn
  transformed data. Lines indicate running CV medians.}
\label{fig:cv-plot}
\end{figure}

@@

<<prepare-msnset-nrom-plot, cache=FALSE, echo=FALSE, keep.fig='none'>>=
p <- plot(normalise(experiment[bsasel], "max"),
          reporters = iTRAQ4, full = FALSE, plot = FALSE)
p <- p + theme_gray(7)
@

\begin{figure}[p]
<<msnexp-norm-plot, dev='pdf', fig.width=5, fig.height=6, echo=FALSE>>=
print(p)
@ 
\caption{Experiment-wide normalised MS2 spectra. The y-axes of the individual 
  spectra is now rescaled between 0 and 1 (highest peak), as opposed to 
  figure \ref{fig:msnexp-plot}. }
\label{fig:msnexp-norm-plot}
\end{figure}

@@

<<combineFeatures,echo=TRUE,cache=FALSE>>=
qnt2 <- combineFeatures(qnt, groupBy = gb, fun = "median")
qnt2
@

@@

<<incompdiss,echo=TRUE,cache=FALSE,keep.source=T>>=
iTRAQ5
incompdiss <- quantify(itraqdata,
                       method = "trap",
                       reporters = iTRAQ5,
                       strict = FALSE,
                       parallel = FALSE,
                       verbose = FALSE)
head(exprs(incompdiss))
@

@@

\begin{figure}[!hbt]
<<incompdiss-plot, dev='pdf', fig.width=5, fig.height=2.5, echo=FALSE, warning=FALSE>>=
dfr <- melt(exprs(incompdiss))
colnames(dfr) <- c("feature", "reporters", "intensity")
dfr$reporters <- sub("iTRAQ5.", "", dfr$reporters)
repsum <- apply(exprs(incompdiss), 1, function(x) sum(x[1:4]))
dfr2 <- data.frame(iTRAQ1to4 = repsum, iTRAQ5 = exprs(incompdiss)[,5])
p1 <- ggplot(data = dfr, aes(x = reporters,y = log10(intensity))) +
  geom_boxplot() +
  theme_gray(6)
p2 <- ggplot(data = dfr2, aes(x = log10(iTRAQ1to4), y = log10(iTRAQ5))) +
  geom_point(alpha = I(1/2)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  stat_smooth(method = lm, se = FALSE) +
  xlab(label = expression(log[10]~sum~114~to~117)) +
  ylab(label = expression(log[10]~145)) +
  theme_gray(6)
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
print(p1, vp = vplayout(1, 1))
print(p2, vp = vplayout(1, 2))
@ 
\caption{Boxplot and scatterplot comparing intensities of the 4 reporter ions 
  (or their sum, on the right) and the incomplete dissociation specific peak. }
\label{fig:incompdiss}
\end{figure}

@@

<<make2exps2, echo=TRUE, cache=FALSE>>=
set.seed(1)
i <- sample(length(itraqdata), 35)
j <- sample(length(itraqdata), 35)
exp1 <- quantify(itraqdata[i], reporters = iTRAQ4,
                 parallel = FALSE, verbose = FALSE)
exp2 <- quantify(itraqdata[j], reporters = iTRAQ4,
                 parallel = FALSE, verbose = FALSE)
table(featureNames(exp1) %in% featureNames(exp2))
exp1 <- combineFeatures(exp1,
                        groupBy = fData(exp1)$ProteinAccession)
exp2 <- combineFeatures(exp2,
                        groupBy = fData(exp2)$ProteinAccession)
head(featureNames(exp1))
head(featureNames(exp1))
@

>>>>

<<environment, cache=FALSE, echo=FALSE>>=
suppressPackageStartupMessages(library("MSnbase"))
suppressPackageStartupMessages(library("zoo"))
if (.Platform$OS.type != "windows")
  suppressPackageStartupMessages(library("doMC"))
suppressPackageStartupMessages(library("multicore"))
library("grid")
library("reshape2")
@

@@

<<readdata, echo=TRUE, cache=FALSE>>=
file <- dir(system.file(package = "MSnbase", dir = "extdata"),
            full.names = TRUE, pattern = "mzXML$")
rawdata <- readMSData(file, msLevel = 2, verbose = FALSE)
@ 
%% $

@@

<<MSnExp, cache=FALSE, echo=TRUE>>=
library("MSnbase")
itraqdata
head(fData(itraqdata))
@

<<experimentsize, echo=FALSE, cache=FALSE>>=
sz <- sum(sapply(assayData(itraqdata), object.size)) + 
  object.size(itraqdata)
sz <- as.numeric(sz)
sz <- round(sz/(1024^2), 2)
@

@@

<<Spectrum, cache=FALSE, echo=TRUE>>=
sp <- itraqdata[["X1"]]
sp
@

@@

<<accessors, cache=FALSE, echo=TRUE>>=
peaksCount(sp)
head(peaksCount(itraqdata))
rtime(sp)
head(rtime(itraqdata))
@

@@

\begin{figure}[!ht]
<<spectrumPlot, dev='pdf', fig.width=4, fig.height=4, fig.keep='high'>>=
plot(sp, reporters = iTRAQ4, full = TRUE) 
@ 
\caption{Raw MS2 spectrum with details about reporter ions.}
\label{fig:spectrum-plot}
\end{figure}

<<bsaSelect, eval=TRUE, echo=FALSE>>=
## bsasel <- fData(itraqdata)$ProteinAccession == "BSA"
bsasel <- 1:3
@ 
%% $

@@

<<subset, echo=TRUE>>=
sel <- fData(itraqdata)$ProteinAccession == "BSA"
bsa <- itraqdata[sel]
bsa
as.character(fData(bsa)$ProteinAccession)
@

@@

\begin{figure}[p]
<<msnexpPlot, dev='pdf', fig.width=3.7, fig.height=4.5, fig.keep='last'>>=
plot(bsa, reporters = iTRAQ4, full = FALSE) + theme_gray(8)
@ 
\caption{Experiment-wide raw MS2 spectra. The y-axes of the individual 
  spectra are automatically rescaled to the same range. 
  See section \ref{sec:norm} to rescale peaks identically. }
\label{fig:msnexp-plot}
\end{figure}

@@

<<removePeaks, echo=TRUE, cache=FALSE>>=
experiment <- removePeaks(itraqdata, t = 400, verbose = FALSE)
## total ion current
ionCount(itraqdata[["X55"]])
ionCount(experiment[["X55"]])
@

@@

<<preprosp, cache=FALSE, echo=FALSE>>=
int <- c(0,1,1,3,1,1,0,0,0,1,3,7,3,1,0)
mz <- c(113.9,114.0,114.05,114.1,114.15,114.2,114.25,
        114.3,114.35,114.4,114.42,114.48,114.5,114.55,114.6)
ppsp <- new("Spectrum2",intensity=int,mz=mz,centroided=FALSE)
p1 <- plot(ppsp, full = TRUE, plot = FALSE) + theme_gray(5) + 
  geom_point(size=3,alpha=I(1/3)) +
  geom_hline(yintercept=3,linetype=2) + 
  ggtitle("Original spectrum") 
p2 <- plot(removePeaks(ppsp,t=3), full=TRUE, plot = FALSE) + 
  theme_gray(5) +
  geom_point(size=3,alpha=I(1/3)) + 
  geom_hline(yintercept=3,linetype=2) + 
  ggtitle("Peaks < 3 removed") 
p3 <- plot(clean(removePeaks(ppsp,t=3)), full = TRUE, plot = FALSE) + 
  theme_gray(5) +
  geom_point(size=3,alpha=I(1/3)) + 
  geom_hline(yintercept=3,linetype=2) + 
  ggtitle("Peaks < 3 removed and cleaned") 
@

\begin{figure}[p]
<<preprocPlot, dev='pdf', fig.width=3, fig.height=6, echo=FALSE>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(3, 1)))
print(p1, vp=vplayout(1,1))
print(p2, vp=vplayout(2,1))
print(p3, vp=vplayout(3,1))
@ 
\caption{This figure illustrated the effect or the \Rfunction{removePeaks} 
  and \Rfunction{clean} methods. The left-most spectrum displays two peaks, 
  of max height 3 and 7 respectively. The middle spectrum shows the result 
  of calling \Rfunction{removePeaks} with argument \Robject{t=3}, which sets 
  all data points of the first peak, whose maximum height is smaller or equal 
  to \Robject{t} to 0. The second peak is unaffected. Calling \Rfunction{clean} 
  after \Rfunction{removePeaks} effectively deletes successive 0 intensities from 
  the spectrum, as shown on the right plot. }
\label{fig:preproc}
\end{figure}

@@

<<trimMz, echo=TRUE, cache=FALSE>>=
range(mz(itraqdata[["X55"]]))
experiment <- trimMz(experiment, mzlim = c(112,120))
range(mz(experiment[["X55"]]))
experiment
@ 
As can be seen above, all processing performed on the experiment is recorded 
and displayed as integral part of the experiment object.

@@

<<reporters, echo=TRUE, cache=FALSE>>=
mz(iTRAQ4)
width(iTRAQ4)
@

@@

\begin{figure}[p]
<<quantitationPlot, dev='pdf',fig.width=5, fig.height=5, echo=FALSE>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(2, 2)))
vplayout <- function(x, y)
  viewport(layout.pos.row = x, layout.pos.col = y)
print(p + ggtitle("Quantitation using 'sum'") +
      geom_point(size = 3, alpha = I(1/3), colour = "red"),
      vp = vplayout(1, 1))
print(p + ggtitle("Quantitation using 'max'") +
      geom_point(aes(x = 114.1, y = 3),
                 alpha = I(1/18),
                 colour = "red", size = 3),
      vp = vplayout(1, 2))
print(p +
      ggtitle("Trapezoidation and strict=FALSE") +
      geom_polygon(alpha = I(1/5), fill = "red"),
      vp = vplayout(2, 1))
print(p +
      ggtitle("Trapezoidation and strict=TRUE") + 
      geom_polygon(aes(x = c(NA, 114.05, 114.05, 114.1, 114.15, 114.15, NA),
                       y=c(NA, 0, 1, 3, 1, 0, NA)), fill = "red", alpha = I(1/5)),
      vp = vplayout(2,2))
@ 
\caption{The different quantitation methods are illustrated above. 
  Quantitation using \Robject{sum} sums all the data points in the peaks to 
  produce, for this example, \Sexpr{quantify(ssp,iTRAQ4[1],method="sum")[[1]]}, 
  whereas method \Robject{max} only uses the peak's maximum intensity, 
  \Sexpr{quantify(ssp,iTRAQ4[1],method="max")[[1]]}. 
  \Robject{Trapezoidation} calculates the area under the peak taking 
  the full with into account (using \Robject{strict=FALSE} gives
  \Sexpr{round(quantify(ssp,iTRAQ4[1],method="trap",strict=FALSE)[[1]],3)}) 
  or only the width as defined by the reporter 
  (using \Robject{strict=TRUE} gives
  \Sexpr{round(quantify(ssp,iTRAQ4[1],method="trap",strict=TRUE)[[1]],3)}). }
\label{fig:quant-methods}
\end{figure}

@@

<<quantify, echo=TRUE, cache=FALSE, keep.source=TRUE>>=
qnt <- quantify(experiment,
                method = "trap",
                reporters = iTRAQ4,
                strict = FALSE,
                parallel = FALSE,
                verbose = FALSE)
qnt
head(exprs(qnt))
@

@@

<<removeNa, echo=TRUE>>=
table(is.na(qnt))
whichRow <- which(is.na((qnt))) %% nrow(qnt)
qnt <- qnt[-whichRow, ]
sum(is.na(exprs(qnt)))
@

@@

<<purityCorrect, echo=TRUE, cache=FALSE>>=
impurities <- matrix(c(0.929, 0.059, 0.002, 0.000,
                       0.020, 0.923, 0.056, 0.001,
                       0.000, 0.030, 0.924, 0.045,
                       0.000, 0.001, 0.040, 0.923),
                     nrow = 4)
qnt.crct <- purityCorrect(qnt, impurities)
head(exprs(qnt))
head(exprs(qnt.crct))
@

@@

<<normalise, echo=TRUE, cache=FALSE>>=
qnt.max <- normalise(qnt, "max")
qnt.sum <- normalise(qnt, "sum")
qnt.quant <- normalise(qnt, "quantiles")
qnt.qrob <- normalise(qnt, "quantiles.robust")
qnt.vsn <- normalise(qnt, "vsn")
@

@@

\begin{figure}[p]
<<normPlot, dev='pdf', fig.width=5, fig.height=7, echo=FALSE>>=
.plot <- function(x,ttl=NULL) {
  boxplot(exprs(x),
          main=ifelse(is.null(ttl),processingData(x)@processing[2],ttl),
          cex.main=.8,
          cex.lab=.5,
          cex.axis=.5,
          cex=.8)
  grid()
}
oldmar <- par()$mar
par(mfrow=c(3,2),mar=c(2.9,2.9,2.9,1))
.plot(qnt, ttl = "Non-normalised data")
.plot(qnt.max, ttl = "Maximum")
.plot(qnt.sum, ttl = "Sum")
.plot(qnt.quant, ttl = "Quantile")
.plot(qnt.qrob, ttl = "Robust quantile")
.plot(qnt.vsn, ttl = "vsn")
@ 
\caption{ Comparison of the normalisation \Robject{MSnSet} methods. 
  Note that vsn also glog-transforms the intensities. }
\label{fig:norm-plot}
\end{figure}

<<cvdata, echo=FALSE, cache=FALSE>>=
sd1 <- apply(log2(exprs(qnt))+10,1,sd)
mn1 <- apply(log2(exprs(qnt))+10,1,mean)
cv1 <- sd1/mn1
sd2 <- apply(exprs(qnt.vsn)+10,1,sd)
mn2 <- apply(exprs(qnt.vsn)+10,1,mean)
cv2 <- sd2/mn2
dfr <- rbind(data.frame(rank=order(mn1),cv=cv1,norm="raw"),
             data.frame(rank=order(mn2),cv=cv2,norm="vsn"))
library("zoo")
## rmed1 <- rollapply(cv1,7,function(x) median(x,na.rm=TRUE))
## rmed2 <- rollapply(cv2,7,function(x) median(x,na.rm=TRUE))
##
## Calling directly rollapply.zoo to make it zoo_1.6-4 compatible.
## The above requires zoo >= 1.7-0, which is as of 15 March 2011
## not yet available on CRAN (only on r-forge).
rmed1 <- zoo:::rollapply.zoo(cv1,7,function(x) median(x,na.rm=TRUE))
rmed2 <- zoo:::rollapply.zoo(cv2,7,function(x) median(x,na.rm=TRUE))
dfr2 <- 
  rbind(data.frame(x=seq(1,70,by=(70/length(rmed1))),
                   y=rmed1,norm="raw"),
        data.frame(x=seq(1,70,by=(70/length(rmed1))),
                   y=rmed2,norm="vsn"))

@@

\begin{figure}[!ht]
<<cvPlot, dev='pdf', fig.width=5, fig.height=4, echo=FALSE>>=
print(p)
@ 
\caption{CV versus signal intensity comparison for log2 and vsn
  transformed data. Lines indicate running CV medians.}
\label{fig:cv-plot}
\end{figure}

@@

<<prepareMsnsetNormPlot, cache=FALSE, echo=FALSE, keep.fig='none'>>=
p <- plot(normalise(experiment[bsasel], "max"),
          reporters = iTRAQ4, full = FALSE, plot = FALSE)
p <- p + theme_gray(7)
@

\begin{figure}[p]
<<msnexpNormPlot, dev='pdf', fig.width=5, fig.height=6, echo=FALSE>>=
print(p)
@ 
\caption{Experiment-wide normalised MS2 spectra. The y-axes of the individual 
  spectra is now rescaled between 0 and 1 (highest peak), as opposed to 
  figure \ref{fig:msnexp-plot}. }
\label{fig:msnexp-norm-plot}
\end{figure}

@@

<<combineFeatures, echo=TRUE, cache=FALSE>>=
qnt2 <- combineFeatures(qnt, groupBy = gb, fun = "median")
qnt2
@

@@

<<incompdiss, echo=TRUE, cache=FALSE, keep.source=TRUE>>=
iTRAQ5
incompdiss <- quantify(itraqdata,
                       method = "trap",
                       reporters = iTRAQ5,
                       strict = FALSE,
                       parallel = FALSE,
                       verbose = FALSE)
head(exprs(incompdiss))
@

@@

\begin{figure}[!hbt]
<<incompdissPlot, dev='pdf', fig.width=5, fig.height=2.5, echo=FALSE, warning=FALSE>>=
dfr <- melt(exprs(incompdiss))
colnames(dfr) <- c("feature", "reporters", "intensity")
dfr$reporters <- sub("iTRAQ5.", "", dfr$reporters)
repsum <- apply(exprs(incompdiss), 1, function(x) sum(x[1:4]))
dfr2 <- data.frame(iTRAQ1to4 = repsum, iTRAQ5 = exprs(incompdiss)[,5])
p1 <- ggplot(data = dfr, aes(x = reporters,y = log10(intensity))) +
  geom_boxplot() +
  theme_gray(6)
p2 <- ggplot(data = dfr2, aes(x = log10(iTRAQ1to4), y = log10(iTRAQ5))) +
  geom_point(alpha = I(1/2)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  stat_smooth(method = lm, se = FALSE) +
  xlab(label = expression(log[10]~sum~114~to~117)) +
  ylab(label = expression(log[10]~145)) +
  theme_gray(6)
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
print(p1, vp = vplayout(1, 1))
print(p2, vp = vplayout(1, 2))
@ 
\caption{Boxplot and scatterplot comparing intensities of the 4 reporter ions 
  (or their sum, on the right) and the incomplete dissociation specific peak. }
\label{fig:incompdiss}
\end{figure}

@@

<<make2exps2, echo=TRUE, cache=FALSE>>=
set.seed(1)
i <- sample(length(itraqdata), 35)
j <- sample(length(itraqdata), 35)
exp1 <- quantify(itraqdata[i], reporters = iTRAQ4,
                 parallel = FALSE, verbose = FALSE)
exp2 <- quantify(itraqdata[j], reporters = iTRAQ4,
                 parallel = FALSE, verbose = FALSE)
table(featureNames(exp1) %in% featureNames(exp2))
exp1 <- combineFeatures(exp1,
                        groupBy = fData(exp1)$ProteinAccession)
exp2 <- combineFeatures(exp2,
                        groupBy = fData(exp2)$ProteinAccession)
head(featureNames(exp1))
head(featureNames(exp2))
@
